Linking hypothesis and number of response options modulate inferred scalar implicature rate

### 2018-05-10 writeup.tex
added abstract and conclusion and final edits'

  \usepackage{graphicx}
  \makeatletter
  \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
  \def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
  \makeatother
  % Scale images if necessary, so that they will not overflow the page
  % margins by default, and it is still possible to overwrite the defaults
  % using explicit options in \includegraphics[width, height, ...]{}
  \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Linking assumptions and number of response options modulate inferred scalar implicature rate},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

% Essential manuscript parts
  \title{Linking assumptions and number of response options modulate inferred
scalar implicature rate}

  \shorttitle{Linking assumptions and implicature rate}

@@

  \authornote{
    Add complete departmental affiliations for each author here. Each new
    line herein must be indented, like this line. Enter author note here.

    Correspondence concerning this article should be addressed to Masoud
    Jasbi, Postal address. E-mail:
    \href{mailto:my@email.com}{\nolinkurl{my@email.com}}
  }

  \abstract{Enter abstract here. Each new line herein must be indented, like this
line.}
  \keywords{scalar implicature; methodology; linking assumption; experimental
pragmatics; truth-value judgment task \\

@@

The past 15 years have seen the rise and development of a bustling and
exciting new field at the intersection of linguistics, psychology, and
philosophy: \emph{experimental pragmatics} (Bott \& Noveck, 2004;
Breheny, Katsos, \& Williams, 2006; Degen \& Tanenhaus, 2015; Geurts \&
Pouscoulous, 2009; Grodner, Klein, Carbary, \& Tanenhaus, 2010; Huang \&
Snedeker, 2009; I. A. Noveck \& Reboul, 2008) \textbf{XXX ADD MORE}.
Experimental pragmatics is devoted to experimentally testing theories of
how language is used in context. How do listeners draw inferences about
the -- often underspecified -- linguistic signal they receive from
speakers? How do speakers choose between the many utterance alternatives
they have at their disposal?

The most prominently studied phenomenon in experimental pragmatics is
undoubtedly \emph{scalar implicature}. Scalar implicatures arise in
virtue of a speaker producing the weaker of two ordered scalemates
(Geurts, 2010; Grice, 1975; Hirschberg, 1985; Horn, 1972). Examples are
provided in (1-2).

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Are scalar inferences default inferences, in the sense that they arise
  unless blocked by (marked) contexts (Degen, 2015; Horn, 1984;
  Levinson, 2000)?
\item
  Are scalar inferences default inferences, in the sense that they are
  computed automatically in online processing and only cancelled by
  context in a second effortful step if required by context) (Bott \&
  Noveck, 2004; Breheny et al., 2006; Degen \& Tanenhaus, 2016; Grodner
  et al., 2010; Huang \& Snedeker, 2009; Politzer-Ahles \& Fiorentino,
  2013; Tomlinson, Bailey, \& Bott, 2013)?
\item
  What are the (linguistic and extra-linguistic) factors that affect
  whether a scalar implicature is derived (Bergen \& Grodner, 2012;
  Bonnefon, Feeney, \& Villejoubert, 2009; Breheny, Ferguson, \& Katsos,
  2013; Breheny et al., 2006; Chemla \& Spector, 2011; De Neys \&
  Schaeken, 2007; Degen, 2015; Degen \& Goodman, 2014; Degen \&
  Tanenhaus, 2015, 2016; Marneffe \& Tonhauser, 2016; Potts, Lassiter,
  Levy, \& Frank, 2015; Zondervan, 2010)?
\item
  How much diversity is there across implicature types, and within
  scalar implicatures across scale types, in whether or not an
  implicature is computed (Doran, Ward, Larson, McNabb, \& Baker, 2012;
  Tiel, Miltenburg, Zevakhina, \& Geurts, 2014)?
\item
  At what age do children acquire the ability to compute implicatures
  (Barner, Brooks, \& Bale, 2011; Horowitz, Schneider, \& Frank, 2017;
  Katsos \& Bishop, 2011; Musolino, 2004; Noveck, 2001; Papafragou \&
  Tantalou, 2004; Stiller, Goodman, \& Frank, 2015)?
\end{enumerate}

@@

A standard measure that has stood proxy for implicature rate across many
studies is the proportion of \enquote{pragmatic} judgments in
truth-value judgment paradigms (Bott \& Noveck, 2004; Chemla \& Spector,
2011; De Neys \& Schaeken, 2007; Degen \& Goodman, 2014; Degen \&
Tanenhaus, 2015; Geurts \& Pouscoulous, 2009; Noveck, 2001; Noveck \&
Posada, 2003). In these kinds of tasks, participants are provided a set
of facts, either presented visually or via their own knowledge of the
world. They are then asked to judge whether a sentence intended to
describe those facts is true or false (or alternatively, whether it is
right or wrong, or they are asked whether they agree or disagree with
the sentence). The crucial condition for assessing implicature rates in
these kinds of studies typically consists of a case where the facts are
such that the stronger alternative is true and the target utterance is
thus also true but underinformative. For instance, Bott and Noveck
(2004) asked participants to judge sentences like \enquote{Some
elephants are mammals}, when world knowledge dictates that all elephants
are mammals. Similarly, Degen and Tanenhaus (2015) asked participants to
judge sentences like \enquote{You got some of the gumballs} in
situations where the visual evidence indicated that the participant
received all the gumballs from a gumball machine. In these kinds of
scenarios, the story goes, if a participant responds \enquote{FALSE},
that indicates that they computed a scalar implicature, eg to the effect
of \enquote{Not all elephants are mammals} or \enquote{You didn't get
all of the gumballs}, which is (globally or contextually) false. If
instead a participant responds \enquote{TRUE}, that is taken to indicate
that they interpreted the utterance literally as `Some, and possibly
all, elephants are mammals' or \enquote{You got some, and possibly all,
of the gumballs}.

Given the centrality of the theoretical notion of \enquote{implicature
rate} to much of experimental pragmatics, there is to date a surprising
lack of discussion of the basic assumption that it is adequately
captured by the proportion of FALSE responses in truth-value judgment
tasks (but see Benz and Gotzner (2014); Geurts and Pouscoulous (2009);
Degen and Goodman (2014); Katsos and Bishop (2011)). Indeed, the scalar
implicature acquisition literature was shaken up when Katsos and Bishop
(2011) showed that simply by introducing an additional response option,
children started looking much more pragmatic than had been previously
observed in a binary judgment paradigm. Katsos and Bishop (2011) allowed
children to distribute 1, 2, or 3 strawberries to a puppet depending on
\enquote{how good the puppet said it}. The result was that children gave
on average fewer strawberries to the puppet when he produced
underinformative utterances compared to when he produced literally true
and pragmatically felicitous utterances, suggesting that children do, in
fact, display pragmatic ability even at ages when they had previously
appeared not to.

But this raises an important question: in truth-value judgment tasks,
how does the researcher know whether an interpretation is literal or the
result of an implicature computation? The binary choice task typically
used is appealing in part because it allows for a direct mapping from
response options -- TRUE and FALSE -- to interpretations -- literal and
pragmatic. That the seeming simplicity of this mapping is illusory
becomes apparent once a third response option is introduced, as in the
Katsos and Bishop (2011) case. How is the researcher to interpret the
intermediate option? Katsos and Bishop (2011) grouped the intermediate
option with the negative endpoint of the scale for the purpose of
categorizing judgments as literal vs.~pragmatic, i.e., they interpreted
the intermediate option as pragmatic. But it seems just as plausible
that they could have grouped it with the positive endpoint of the scale
and taken the hard line that only truly FALSE responses constitute
evidence of a full-fledged implicature. The point here is that there has
been remarkably little consideration of \emph{linking functions} between
behavioral measures and theoretical constructs in experimental
pragmatics, a problem in many subfields of psycholinguistics (Tanenhaus,
2004). We argue that it is time to engage more seriously with these
issues.

We begin by reporting an experiment that addresses the following
question: do the number of response options provided in a truth-value
judgment task and the way that responses are grouped into pragmatic
(\enquote{SI}) and literal (\enquote{no SI}) change inferences about
scalar implicature rates? Note that this way of asking the question
presupposes two things: first, that whatever participants are doing in a
truth-value judgment task, the behavioral measure can be interpreted as
providing a measure of \emph{interpretation}. And second, that listeners
either do or do not compute an implicature on any given occasion. In the
General Discussion we will discuss both of these issues. First,
following Degen and Goodman (2014), we will offer some remarks on why
truth-value judgment tasks are better thought of as measuring
participants' estimates of speakers' \emph{production} probabilities.
This will suggest a completely different class of linking functions. And
second, we discuss an alternative conception of scalar implicature as a
probabilistic phenomeonen, a view that has recently rose to prominence
in the subfield of probabilistic pragmatics (Franke \& Jäger, 2016;
Goodman \& Frank, 2016). This alternative conception of scalar
implicature, we argue, affords developing and testing quantitative
linking functions in a rigorous and motivated way.

Consider a setup in which a listener is presented a card with a
depiction of either one or two animals (see Figure
\ref{fig:linkvisualization} for an example). As in a standard
truth-value judgment task, the listener then observes an
underinformative utterance about this card (e.g., \enquote{There is a
cat or a dog on the card}) and is asked to provide a judgment on a scale
with 2, 3, 4, or 5 response options, with endpoints \enquote{wrong} and
\enquote{right.}\footnote{An open question concerns the extent to which
  the labeling of points on the scale affects judgments (e.g.,
  \enquote{wrong}--\enquote{right} vs. \enquote{false}--\enquote{true}
  vs. \enquote{disagree}--\enquote{agree}). While some studies have used
  \enquote{false}--\enquote{true}, others have argued that judging truth
  may lead to meta-linguistic reasoning in participants that could
  distort judgments.} In the binary case, this reproduces the standard
truth-value judgment task. Figure \ref{fig:linkvisualization}
exemplifies (some of) the researcher's options for grouping responses.
Under what we will call the \enquote{Strong link} assumption, only the
negative endpoint of the scale is interpreted as evidence for a scalar
implicature having been computed. Under the \enquote{Weak link}
assumption, in contrast, any response that does not correspond to the
positive endpoint of the scale is interpreted as evidence for a scalar
implicature having been computed. Intermediate grouping schemes are also
possible, but these are the ones we will consider here. Note that for
the binary case, the Weak and Strong link return the same categorization
scheme, but for any number of response options greater than 2, the Weak
and Strong link can in principle lead to differences in inferences about
implicature rate.

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/linkvisualization-1.pdf}
\caption{\label{fig:linkvisualization}Strong and weak link from response
options to researcher inference about scalar implicature rate,
exemplified for the disjunctive utterance when the conjunction is true.}
\end{figure}

Let's examine an example. Assume three response options (wrong, neither,
right). Assume further that a third of participants each gave each of
the three responses, i.e., the distributions of responses is 1/3, 1/3,
and 1/3. Under the Strong link, we infer that this task yielded an
implicature rate of 2/3. Under the Weak link, we infer that this task
yielded an implicature rate of 1/3. This is quite a drastic difference
if we are for instance interested in whether scalar implicatures are
inference defaults and we would like to interpret an implicature rate of
above an arbitrary threshold (e.g., 50\%) as evidence for such a claim.
Under the Strong link, we would conclude that scalar implicatures are
not defaults. Under the Weak link, we would conclude that they are. In
the experiment reported in the following section, we presented
participants with exactly this setup. We manipulated the number of
response options between participants and analyzed the results under
different linking assumptions.

@@

Participants played an online card game in which they were asked to
judge descriptions of the contents of cards. Different groups of
participants were presented with different numbers of response options.
On critical trials, participants were presented with descriptions for
the cards that typically result in exhaustivity implicatures
(\enquote{There is a cat on the card} when there was a cat and a dog) or
scalar implicatures (\enquote{There is a cat or a dog on the card} when
there was a cat and a dog). We categorized their responses on such
trials according to the weak and the strong link introduced above, and
tested whether the number of response options and the linking
assumptions led to different conclusions about the rate of computed
implicatures in the experimental task.

@@

200 participants were recruited via Amazon Mechanical Turk. They
optionally provided demographic information at the end of the study.
Participants' mean age was 35. We also asked participants if they had
any prior training in logic. 40 participants reported that they did,
while 160 had no prior training in logic. \textbf{JD: Do we do any
analysis of logic training???} All participants' data was included in
the final analysis.

@@

The study was administered online through Amazon Mechanical
Turk.\footnote{The experiment can be viewed
  \href{https://cdn.rawgit.com/thegricean/si-paradigms/94a590f0/experiments/main/1_methods/online_experiment/connective_game.html}{here}.}
Participants were first introduced to the set of cards we used in the
study (Figure \ref{fig:stimuli}). Each card depicted one or two animals,
where an animal could be either a cat, a dog, or an elephant. Then
participants were introduced to a blindfolded fictional character called
Bob. Bob was blindfolded to avoid violations of ignorance expectations
associated with the use of disjunction ({\textbf{???}}; Chierchia et
al., 2001) \textbf{JD: Masoud, can you insert citation pls? Geurts,
Sauerland?}. Participants were told that Bob would guess the contents of
the cards and their task was to indicate whether Bob's guess was wrong
or right. On each trial, participants saw a card and a sentence
representing Bob's guess. For example, they saw a card with a cat and
read the sentence \enquote{There is a cat on the card.} They then
provided an assessment of Bob's guess. The study ended after 24 trials.

@@

\subsection{Results}\label{results}

@@

In the binary task, participants were at or close to ceiling in
responding \enquote{right} and \enquote{wrong} on unambiguously true and
false trials, respectively (see Figure \ref{fig:binaryPlot}). However,
on underinformative trials (\enquote{cat} and \enquote{cat or dog}
description for card with cat and dog), we observe pragmatic behavior:
on exhaustive trials, participants judged the utterance \enquote{wrong}
14\% of the time; on scalar trials, participants judged the utterance
\enquote{wrong} 38\% of the time. That is, both under the weak and the
strong link assumptions introduced in the Introduction, inferred
implicature rate on exhaustive trials is 14\% and on scalar trials 38\%.

@@

In the ternary task, participants were also at or close to ceiling in
responding \enquote{right} and \enquote{wrong} on unambiguously true and
false trials, respectively (see Figure \ref{fig:ternaryPlot}). And
again, on underinformative trials (\enquote{cat} and \enquote{cat or
dog} description for card with cat and dog), we observed pragmatic
behavior: on exhaustive trials, participants considered the guess
\enquote{wrong} 8\% of the time and neither wrong nor right 12\% of the
time. On scalar trials, participants judged the guess \enquote{wrong}
23\% of the time and \enquote{neither} 11\% of the time. This means that
the weak and strong link lead to different conclusions about implicature
rates on the ternary task. Under the weak link, inferred implicature
rate on exhaustive trials is 20\%; under the strong link it is only 8\%.
Similarly, under the weak link, inferred implicature rate on scalar
trials is 34\%; under the strong link it is only 23\%.

@@

In the quaternary task (Figure \ref{fig:quaternaryPlot}), participants
were again at or close to ceiling in responding \enquote{right} and
\enquote{wrong} on 4 of the 6 unambiguously true and false trials.
However, with four response options, two of the control conditions
appear to be showing signs of pragmatic infelicity: when a conjunction
was used and only one of the animals was on the card, participants
considered the guess \enquote{wrong} most of the time, but they often
considered it \enquote{kinda wrong} or even \enquote{kinda right}. This
suggests that perhaps participants considered a notion of partially true
or correct statement in our experimental setting. Disjunctive
descriptions of cards with only one animal, while previously at ceiling
for \enquote{right} responses, were downgraded to only \enquote{kinda
right} 26\% of the time, presumably because these utterances are also
underinformative, though the degree of underinformativeness may be less
egregious than on scalar trials.

@@

Thus, we are again forced to draw different conclusions about
implicature rates depending on whether we assume the weak link or the
strong link. Under the weak link, inferred implicature rate on
exhaustive trials is 73\%; under the strong link it is only 2\%.
Similarly, under the weak link, inferred implicature rate on scalar
trials is 61\%; under the strong link it is only 6\%.

@@

Finally, Figure \ref{fig:quinaryPlot} shows the proportion of responses
in the quinary task. Performance on the 4 pragmatically felicitous
control trials was again at floor and ceiling, respectively. The 2
control conditions that the quaternary task had revealed pragmatic
infelicity in again displayed that pragmatic infelicity in the quinary
task, suggesting that this is a robust type of pragmatic infelicity
that, however, requires fine-grained enough response options to be
detected experimentally.

@@

Thus, we would again draw different conclusions about implicature rates
depending on whether we assume the weak link or the strong link. Under
the weak link, inferred implicature rate on exhaustive trials is 76 and
76\%; under the strong link it is only 2\%. Similarly, under the weak
link, inferred implicature rate on scalar trials is 63\%; under the
strong link it is only 6\%.

>>>>

  \usepackage{graphicx}
  \makeatletter
  \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
  \def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
  \makeatother
  % Scale images if necessary, so that they will not overflow the page
  % margins by default, and it is still possible to overwrite the defaults
  % using explicit options in \includegraphics[width, height, ...]{}
  \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Linking hypothesis and number of response options modulate inferred scalar implicature rate},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

% Essential manuscript parts
  \title{Linking hypothesis and number of response options modulate inferred
scalar implicature rate}

  \shorttitle{Linking hypotheses and implicature rate}

@@

  \authornote{
    XXX Add complete departmental affiliations for each author here. Each
    new line herein must be indented, like this line. Enter author note
    here.

    Correspondence concerning this article should be addressed to Masoud
    Jasbi, XXX Postal address. E-mail: XXX
    \href{mailto:my@email.com}{\nolinkurl{my@email.com}}
  }

  \abstract{The past 15 years have seen increasing experimental investigations of
core pragmatic questions in the ever more active and lively field of
experimental pragmatics. Within experimental pragmatics, many of the
core questions have relied on the operationalization of the theoretical
notion of `implicature rate'. Implicature rate based results have
informed the work on acquisition, online processing, and scalar
diversity, inter alia. Despite its theoretical importance. Implicature
rate has typically been quantified as the proportion of `pragmatic'
judgments in two-alternative forced choice truth value judgment tasks.
Despite its theoretical importance, this linking hypothesis from
implicature rate to behavioral responses has never been extensively
tested. Here we show that two factors dramatically affect the
`implicature rate' inferred from truth value judgment tasks: a) the
number of responses provided to participants; and b) the linking
hypothesis about what constitutes a `pragmatic' judgment. We argue that
it is time for the field of experimental pragmatics to engage more
seriously with its foundational assumptions about how theoretical
notions map onto behaviorally measurable quantities, and present a
sketch of an alternative linking hypothesis that derives behavior in
truth value judgment tasks from probabilistic utterance expectations.}
  \keywords{scalar implicature; methodology; linking hypothesis; experimental
pragmatics; truth value judgment task \\

@@

The past 15 years have seen the rise and development of a bustling and
exciting new field at the intersection of linguistics, psychology, and
philosophy: \emph{experimental pragmatics} (Barner, Brooks, \& Bale,
2011; Bonnefon, Feeney, \& Villejoubert, 2009; Bott \& Chemla, 2016;
Bott \& Noveck, 2004; Breheny, Ferguson, \& Katsos, 2013; Breheny,
Katsos, \& Williams, 2006; Chierchia et al., 2001; De Neys \& Schaeken,
2007; Degen \& Tanenhaus, 2015, 2016; Geurts \& Pouscoulous, 2009;
Grodner, Klein, Carbary, \& Tanenhaus, 2010; Huang \& Snedeker, 2009;
Katsos \& Bishop, 2011; I. A. Noveck \& Reboul, 2008; Noveck \& Posada,
2003; Papafragou \& Tantalou, 2004; Tiel, Miltenburg, Zevakhina, \&
Geurts, 2014; Tomlinson, Bailey, \& Bott, 2013). Experimental pragmatics
is devoted to experimentally testing theories of how language is used in
context. How do listeners draw inferences about the -- often
underspecified -- linguistic signal they receive from speakers? How do
speakers choose between the many utterance alternatives they have at
their disposal?

The most prominently studied phenomenon in experimental pragmatics is
undoubtedly \emph{scalar implicature}. Scalar implicatures arise as a
result of a speaker producing the weaker of two ordered scalemates
(Geurts, 2010; Grice, 1975; Hirschberg, 1985; Horn, 1972). Examples are
provided in (1-2).

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Are scalar inferences default inferences, in the sense that they arise
  unless blocked by (marked) contexts (Degen, 2015; Horn, 1984;
  Levinson, 2000)?
\item
  Are scalar inferences default inferences, in the sense that they are
  computed automatically in online processing and only cancelled by
  context in a second effortful step if required by context) (Bott \&
  Noveck, 2004; Breheny et al., 2006; Degen \& Tanenhaus, 2016; Grodner
  et al., 2010; Huang \& Snedeker, 2009; Politzer-Ahles \& Fiorentino,
  2013; Tomlinson et al., 2013)?
\item
  What are the (linguistic and extra-linguistic) factors that affect
  whether a scalar implicature is derived (Bergen \& Grodner, 2012;
  Bonnefon et al., 2009; Breheny et al., 2013, 2006; Chemla \& Spector,
  2011; De Neys \& Schaeken, 2007; Degen, 2015; Degen \& Goodman, 2014;
  Degen \& Tanenhaus, 2015, 2016; Marneffe \& Tonhauser, 2016; Potts,
  Lassiter, Levy, \& Frank, 2015; Zondervan, 2010)?
\item
  How much diversity is there across implicature types, and within
  scalar implicatures across scale types, in whether or not an
  implicature is computed (Doran, Ward, Larson, McNabb, \& Baker, 2012;
  Tiel et al., 2014)?
\item
  At what age do children acquire the ability to compute implicatures
  (Barner et al., 2011; Horowitz, Schneider, \& Frank, 2017; Katsos \&
  Bishop, 2011; Musolino, 2004; Noveck, 2001; Papafragou \& Tantalou,
  2004; Stiller, Goodman, \& Frank, 2015)?
\end{enumerate}

@@

A standard measure that has stood as a proxy for implicature rate across
many studies is the proportion of \enquote{pragmatic} judgments in truth
value judgment paradigms (Bott \& Noveck, 2004; Chemla \& Spector, 2011;
De Neys \& Schaeken, 2007; Degen \& Goodman, 2014; Degen \& Tanenhaus,
2015; Geurts \& Pouscoulous, 2009; Noveck, 2001; Noveck \& Posada,
2003). In these kinds of tasks, participants are provided a set of
facts, either presented visually or via their own knowledge of the
world. They are then asked to judge whether a sentence intended to
describe those facts is true or false (or alternatively, whether it is
right or wrong, or they are asked whether they agree or disagree with
the sentence). The crucial condition for assessing implicature rates in
these kinds of studies typically consists of a case where the facts are
such that the stronger alternative is true and the target utterance is
thus also true but underinformative. For instance, Bott and Noveck
(2004) asked participants to judge sentences like \enquote{Some
elephants are mammals}, when world knowledge dictates that all elephants
are mammals. Similarly, Degen and Tanenhaus (2015) asked participants to
judge sentences like \enquote{You got some of the gumballs} in
situations where the visual evidence indicated that the participant
received all the gumballs from a gumball machine. In these kinds of
scenarios, the story goes, if a participant responds \enquote{FALSE},
that indicates that they computed a scalar implicature, eg to the effect
of \enquote{Not all elephants are mammals} or \enquote{You didn't get
all of the gumballs}, which is (globally or contextually) false. If
instead a participant responds \enquote{TRUE}, that is taken to indicate
that they interpreted the utterance literally as `Some, and possibly
all, elephants are mammals' or \enquote{You got some, and possibly all,
of the gumballs}.

Given the centrality of the theoretical notion of \enquote{implicature
rate} to much of experimental pragmatics, there is to date a surprising
lack of discussion of the basic assumption that it is adequately
captured by the proportion of FALSE responses in truth value judgment
tasks (but see Benz and Gotzner (2014); Geurts and Pouscoulous (2009);
Degen and Goodman (2014); Katsos and Bishop (2011)). Indeed, the scalar
implicature acquisition literature was shaken up when Katsos and Bishop
(2011) showed that simply by introducing an additional response option,
children started looking much more pragmatic than had been previously
observed in a binary judgment paradigm. Katsos and Bishop (2011) allowed
children to distribute 1, 2, or 3 strawberries to a puppet depending on
\enquote{how good the puppet said it}. The result was that children gave
on average fewer strawberries to the puppet when he produced
underinformative utterances compared to when he produced literally true
and pragmatically felicitous utterances, suggesting that children do, in
fact, display pragmatic ability even at ages when they had previously
appeared not to.

But this raises an important question: in truth value judgment tasks,
how does the researcher know whether an interpretation is literal or the
result of an implicature computation? The binary choice task typically
used is appealing in part because it allows for a direct mapping from
response options -- TRUE and FALSE -- to interpretations -- literal and
pragmatic. That the seeming simplicity of this mapping is illusory
becomes apparent once a third response option is introduced, as in the
Katsos and Bishop (2011) case. How is the researcher to interpret the
intermediate option? Katsos and Bishop (2011) grouped the intermediate
option with the negative endpoint of the scale for the purpose of
categorizing judgments as literal vs.~pragmatic, i.e., they interpreted
the intermediate option as pragmatic. But it seems just as plausible
that they could have grouped it with the positive endpoint of the scale
and taken the hard line that only truly FALSE responses constitute
evidence of a full-fledged implicature. The point here is that there has
been remarkably little consideration of \emph{linking hypothesiss}
between behavioral measures and theoretical constructs in experimental
pragmatics, a problem in many subfields of psycholinguistics (Tanenhaus,
2004). We argue that it is time to engage more seriously with these
issues.

We begin by reporting an experiment that addresses the following
question: do the number of response options provided in a truth value
judgment task and the way that responses are grouped into pragmatic
(\enquote{SI}) and literal (\enquote{no SI}) change inferences about
scalar implicature rates? Note that this way of asking the question
presupposes two things: first, that whatever participants are doing in a
truth value judgment task, the behavioral measure can be interpreted as
providing a measure of interpretation; and second, that listeners either
do or do not compute an implicature on any given occasion. In the
General Discussion we will discuss both of these issues. First,
following Degen and Goodman (2014), we will offer some remarks on why
truth value judgment tasks are better thought of as measuring
participants' estimates of speakers' \emph{production} probabilities.
This will suggest a completely different class of linking hypothesiss.
Next, we discuss an alternative conception of scalar implicature as a
probabilistic phenomeonen, a view that has recently rose to prominence
in the subfield of probabilistic pragmatics (Franke \& Jäger, 2016;
Goodman \& Frank, 2016). This alternative conception of scalar
implicature, we argue, affords developing and testing quantitative
linking hypothesiss in a rigorous and motivated way.

Consider a setup in which a listener is presented a card with a
depiction of either one or two animals (see Figure
\ref{fig:linkvisualization} for an example). As in a standard truth
value judgment task, the listener then observes an underinformative
utterance about this card (e.g., \enquote{There is a cat or a dog on the
card}) and is asked to provide a judgment on a scale with 2, 3, 4, or 5
response options, with endpoints \enquote{wrong} and
\enquote{right.}\footnote{An open question concerns the extent to which
  the labeling of points on the scale affects judgments (e.g.,
  \enquote{wrong}--\enquote{right} vs. \enquote{false}--\enquote{true}
  vs. \enquote{disagree}--\enquote{agree}). While some studies have used
  \enquote{false}--\enquote{true}, others have argued that judging truth
  may lead to meta-linguistic reasoning in participants that could
  distort judgments.} In the binary case, this reproduces the standard
truth value judgment task. Figure \ref{fig:linkvisualization}
exemplifies (some of) the researcher's options for grouping responses.
Under what we will call the \enquote{Strong link} assumption, only the
negative endpoint of the scale is interpreted as evidence for a scalar
implicature having been computed. Under the \enquote{Weak link}
assumption, in contrast, any response that does not correspond to the
positive endpoint of the scale is interpreted as evidence for a scalar
implicature having been computed. Intermediate grouping schemes are also
possible, but these are the ones we will consider here. Note that for
the binary case, the Weak and Strong link return the same categorization
scheme, but for any number of response options greater than 2, the Weak
and Strong link can in principle lead to differences in inferences about
implicature rate.

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/linkvisualization-1.pdf}
\caption{\label{fig:linkvisualization}Strong and Weak link from response
options to researcher inference about scalar implicature rate,
exemplified for the disjunctive utterance when the conjunction is true.}
\end{figure}

Let's examine an example. Assume three response options (wrong, neither,
right). Assume further that each of the three responses was selected by
a third of participants, i.e., the distributions of responses is 1/3,
1/3, and 1/3. Under the Strong link, we infer that this task yielded an
implicature rate of 2/3. Under the Weak link, we infer that this task
yielded an implicature rate of 1/3. This is quite a drastic difference
if we are, for instance, interested in whether scalar implicatures are
inference defaults and we would like to interpret an implicature rate of
above an arbitrary threshold (e.g., 50\%) as evidence for such a claim.
Under the Strong link, we would conclude that scalar implicatures are
not defaults. Under the Weak link, we would conclude that they are. In
the experiment reported in the following section, we presented
participants with exactly this setup. We manipulated the number of
response options between participants and analyzed the results under
different linking hypothesiss.

@@

Participants played an online card game in which they were asked to
judge descriptions of the contents of cards. Different groups of
participants were presented with different numbers of response options.
On critical trials, participants were presented with descriptions for
the cards that typically result in exhaustivity implicatures
(\enquote{There is a cat on the card} when there was a cat and a dog) or
scalar implicatures (\enquote{There is a cat or a dog on the card} when
there was a cat and a dog). We categorized their responses on such
trials according to the weak and the Strong link introduced above, and
tested whether the number of response options and the linking
hypothesiss led to different conclusions about the rate of computed
implicatures in the experimental task.

@@

200 participants were recruited via Amazon Mechanical Turk. They
optionally provided demographic information at the end of the study.
Participants' mean age was 35. We also asked participants if they had
any prior training in logic. 40 participants reported that they did,
while 160 had no prior training in logic. All participants' data was
included in the final analysis.

@@

The study was administered online through Amazon Mechanical
Turk.\footnote{The experiment can be viewed
  \href{https://cdn.rawgit.com/thegricean/si-paradigms/94a590f0/experiments/main/1_methods/online_experiment/connective_game.html}{here}.}
Participants were first introduced to the set of cards we used in the
study (Figure \ref{fig:stimuli}). Each card depicted one or two animals,
where an animal could be either a cat, a dog, or an elephant. Then
participants were introduced to a blindfolded fictional character called
Bob. Bob was blindfolded to avoid violations of ignorance expectations
associated with the use of disjunction ({\textbf{???}}; Chierchia et
al., 2001) Participants were told that Bob would guess the contents of
the cards and their task was to indicate whether Bob's guess was wrong
or right. On each trial, participants saw a card and a sentence
representing Bob's guess. For example, they saw a card with a cat and
read the sentence \enquote{There is a cat on the card.} They then
provided an assessment of Bob's guess. The study ended after 24 trials.

@@

\subsection{Results and discussion}\label{results-and-discussion}

@@

In the binary task, participants were at or close to ceiling in
responding \enquote{right} and \enquote{wrong} on unambiguously true and
false trials, respectively (see Figure \ref{fig:binaryPlot}). However,
on underinformative trials (i.e.~a \enquote{cat} or \enquote{cat or dog}
description for a card with both a cat and a dog), we observe pragmatic
behavior: on exhaustive trials, participants judged the utterance
\enquote{wrong} 14\% of the time; on scalar trials, participants judged
the utterance \enquote{wrong} 38\% of the time. That is, both under the
Weak and Strong link assumptions introduced in the Introduction,
inferred implicature rate on exhaustive trials is 14\% and on scalar
trials 38\%.

@@

In the ternary task, participants were also at or close to ceiling in
responding \enquote{right} and \enquote{wrong} on unambiguously true and
false trials, respectively (see Figure \ref{fig:ternaryPlot}). And
again, on underinformative trials (a \enquote{cat} and \enquote{cat or
dog} description for a card with both a cat and a dog), we observed
pragmatic behavior: on exhaustive trials, participants considered the
guess \enquote{wrong} 8\% of the time and neither wrong nor right 12\%
of the time. On scalar trials, participants judged the guess
\enquote{wrong} 23\% of the time and \enquote{neither} 11\% of the time.
This means that the Weak and Strong link lead to different conclusions
about implicature rates on the ternary task. Under the Weak link,
inferred implicature rate on exhaustive trials is 20\%; under the Strong
link it is only 8\%. Similarly, under the Weak link, inferred
implicature rate on scalar trials is 34\%; under the Strong link it is
only 23\%.

@@

In the quaternary task (Figure \ref{fig:quaternaryPlot}), participants
were again at or close to ceiling in responding \enquote{right} and
\enquote{wrong} on 4 of the 6 unambiguously true and false trials.
However, with four response options, two of the control conditions
appear to be showing signs of pragmatic infelicity: when a conjunction
was used and only one of the animals was on the card, participants
considered the guess \enquote{wrong} most of the time, but they often
considered it \enquote{kinda wrong} or even \enquote{kinda right}. This
suggests that perhaps participants considered the notion of a partially
true or correct statement in our experimental setting. Disjunctive
descriptions of cards with only one animal, while previously at ceiling
for \enquote{right} responses, were downgraded to only \enquote{kinda
right} 26\% of the time, presumably because these utterances are also
underinformative, though the degree of underinformativeness may be less
egregious than on scalar trials.

@@

Thus, we are again forced to draw different conclusions about
implicature rates depending on whether we assume the Weak link or the
Strong link. Under the Weak link, inferred implicature rate on
exhaustive trials is 73\%; under the Strong link it is only 2\%.
Similarly, under the Weak link, inferred implicature rate on scalar
trials is 61\%; under the Strong link it is only 6\%.

@@

Finally, Figure \ref{fig:quinaryPlot} shows the proportion of responses
in the quinary task. Performance on the 4 pragmatically felicitous
control trials was again at floor and ceiling, respectively. The 2
control conditions in which the quaternary task had revealed pragmatic
infelicity again displayed that pragmatic infelicity in the quinary
task, suggesting that this is a robust type of pragmatic infelicity
that, nonetheless, requires fine-grained enough response options to be
detected experimentally.

@@

Thus, we would again draw different conclusions about implicature rates
depending on whether we assume the Weak link or the Strong link. Under
the Weak link, inferred implicature rate on exhaustive trials is 76 and
76\%; under the Strong link it is only 2\%. Similarly, under the Weak
link, inferred implicature rate on scalar trials is 63\%; under the
Strong link it is only 6\%.
